Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics

### 2018-12-13 scagnostics.Rmd
Fixed index sorting in trace plots.

```{r pipeFirstRun, fig.width=7, fig.height=7, out.width = "0.8\\textwidth", fig.cap="Scouting for the pipe using optimization method 2 on the TIC index. Other PPI values also shown for interest. Only the values for the target plane are shown because the optimization is done off-line. The TIC maximum value for this data is small,  even though it identifies the pipe. The holes, convex, skinny, and to some extent MIC, mirror the performance of TIC for this structure."}
if(!file.exists("cache/pipeFirstRun.rda")){
  pipeTourRes <- getProj(pipe1000, pipeTourFull, "Pipe", 1000)
  #pipeTourResMelt <- melt(pipeTourRes, id=c("t", "name", "size"))
  #never used?
  #save(pipeTourRes, pipeTourResMelt, file = "cache/pipeFirstRun.rda")
  save(pipeTourRes, file = "cache/pipeFirstRun.rda")
} else {
  load("cache/pipeFirstRun.rda")
}
newBasisPipeTour <- pipeTourRes %>%
  mutate(convex = 1-convex) %>% 
  rowwise() %>%
  mutate(holes = rescaleHoles(holes)) %>%
  select(-cmass) %>%
  gather(variable, value, -t, -name, -size) %>%
  filter(t %in% which(attributes(interpolate(pipeTour))$new_basis))
ggplot(newBasisPipeTour, aes(x=t, y=value)) +
  geom_point(mapping=aes(color=(variable=="TIC"))) + 
  geom_line(mapping=aes(color=(variable=="TIC"))) + 
  scale_colour_manual(values = c("black", "red")) +
  facet_wrap(~variable, ncol=1, strip.position = "right") +
  guides(color=FALSE) +
  ylim(c(0,1)) +
  xlab("Number of projections") +
  ylab("PPI value") 
```

@@

```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising  the splines2d index, using method 3, for the sine data, with $n=1000$. Anchor planes are marked by the blue vertical lines, and are closer to each other approaching the maxima. The sine is found relatively easily, by splines2D, and it si indicated that MIC, TIC, dcor2D and convex would also likely find this structure.", results="hide"}
set.seed(2018)
if(!file.exists("cache/findsine.rda")){
  sineResc <- as.matrix(sin1000)
  sineTour <- save_history(sineResc,
                          guided_tour(splineIndex()))
  sineTourFull <- as.list(interpolate(sineTour))
  sineTourRes <- getProj(sin1000, sineTourFull, "Sine", 1000)
  #sineTourResMelt <- melt(sineTourRes, id=c("t", "name", "size"))
  #save(sineResc, sineTour, sineTourFull, sineTourRes, sineTourResMelt, file = "cache/findsine.rda")
  save(sineResc, sineTour, sineTourFull, sineTourRes, file = "cache/findsine.rda")
} else {
  load("cache/findsine.rda")
}
anchorIdx <- as.tibble(which(attributes(interpolate(sineTour))$new_basis))
sineTourResMelt <-sineTourRes %>%
  mutate(convex = 1-convex) %>%
  rowwise() %>%
  mutate(holes = rescaleHoles(holes)) %>%
  select(-cmass) %>%
  gather(variable, value, -t, -name, -size)
ggplot(sineTourResMelt, aes(x=t, y=value)) +
  geom_line(mapping=aes(color=(variable=="splines2d"))) + 
  facet_wrap(~variable, ncol=1, scales = "free_y") +
  scale_colour_manual(values = c("black", "red")) +
  facet_wrap(~variable, ncol=1) +
  geom_vline(data=anchorIdx, mapping=aes(xintercept=value), color="blue", alpha=0.3) +
  guides(color=FALSE) +
  ylim(c(0,1)) + xlab("Number of projections") + ylab("PPI value")
```

>>>>

```{r pipeFirstRun, fig.width=7, fig.height=7, out.width = "0.8\\textwidth", fig.cap="Scouting for the pipe using optimization method 2 on the TIC index. Other PPI values also shown for interest. Only the values for the target plane are shown because the optimization is done off-line. The TIC maximum value for this data is small,  even though it identifies the pipe. The holes, convex, skinny, and to some extent MIC, mirror the performance of TIC for this structure."}
if(!file.exists("cache/pipeFirstRun.rda")){
  pipeTourRes <- getProj(pipe1000, pipeTourFull, "Pipe", 1000)
  #pipeTourResMelt <- melt(pipeTourRes, id=c("t", "name", "size"))
  #never used?
  #save(pipeTourRes, pipeTourResMelt, file = "cache/pipeFirstRun.rda")
  save(pipeTourRes, file = "cache/pipeFirstRun.rda")
} else {
  load("cache/pipeFirstRun.rda")
}
newBasisPipeTour <- pipeTourRes %>%
  mutate(convex = 1-convex) %>% 
  rowwise() %>%
  mutate(holes = rescaleHoles(holes)) %>%
  select(-cmass) %>%
  gather(variable, value, -t, -name, -size) %>%
  filter(t %in% which(attributes(interpolate(pipeTour))$new_basis)) %>%
  mutate(variable = fct_relevel(variable, "holes", "convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))

ggplot(newBasisPipeTour, aes(x=t, y=value)) +
  geom_point(mapping=aes(color=(variable=="TIC"))) + 
  geom_line(mapping=aes(color=(variable=="TIC"))) + 
  scale_colour_manual(values = c("black", "red")) +
  facet_wrap(~variable, ncol=1, strip.position = "right") +
  guides(color=FALSE) +
  ylim(c(0,1)) +
  xlab("Number of projections") +
  ylab("PPI value") 
```

@@

```{r findsine, fig.width=6, fig.height=7, out.width = "0.8\\textwidth", fig.cap="Guided tour optimising  the splines2d index, using method 3, for the sine data, with $n=1000$. Anchor planes are marked by the blue vertical lines, and are closer to each other approaching the maxima. The sine is found relatively easily, by splines2D, and it si indicated that MIC, TIC, dcor2D and convex would also likely find this structure.", results="hide"}
set.seed(2018)
if(!file.exists("cache/findsine.rda")){
  sineResc <- as.matrix(sin1000)
  sineTour <- save_history(sineResc,
                          guided_tour(splineIndex()))
  sineTourFull <- as.list(interpolate(sineTour))
  sineTourRes <- getProj(sin1000, sineTourFull, "Sine", 1000)
  #sineTourResMelt <- melt(sineTourRes, id=c("t", "name", "size"))
  #save(sineResc, sineTour, sineTourFull, sineTourRes, sineTourResMelt, file = "cache/findsine.rda")
  save(sineResc, sineTour, sineTourFull, sineTourRes, file = "cache/findsine.rda")
} else {
  load("cache/findsine.rda")
}
anchorIdx <- as.tibble(which(attributes(interpolate(sineTour))$new_basis))
sineTourResMelt <-sineTourRes %>%
  mutate(convex = 1-convex) %>%
  rowwise() %>%
  mutate(holes = rescaleHoles(holes)) %>%
  select(-cmass) %>%
  gather(variable, value, -t, -name, -size) %>%
  mutate(variable = fct_relevel(variable, "holes", "convex", "skinny", "stringy", "dcor2d", "splines2d", "MIC", "TIC"))
ggplot(sineTourResMelt, aes(x=t, y=value)) +
  geom_line(mapping=aes(color=(variable=="splines2d"))) + 
  facet_wrap(~variable, ncol=1, scales = "free_y") +
  scale_colour_manual(values = c("black", "red")) +
  facet_wrap(~variable, ncol=1, strip.position = "right") +
  geom_vline(data=anchorIdx, mapping=aes(xintercept=value), color="blue", alpha=0.3) +
  guides(color=FALSE) +
  ylim(c(0,1)) + xlab("Number of projections") + ylab("PPI value")
```
